---
title: "Warm-Up Code"
output: html_document
---

### Goal: Observe the syntax and describe what each code example does. Try to predict the output. Then run the code to check your understanding.

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = "hide")
```

**Single tests with `if... else`**

Example 1
```{r}
age <- 14
if (age > 16) {
  print("You can get a driving license")
} else {
  print("You cannot drive")
}
```

Example 2
```{r}
movie_rating <- c(4.5, 4.2, 5.0, 3.8, 2.9, 1.7)

movie_avg_rating <- mean(movie_rating)

if (movie_avg_rating >= 4) {
  print("This movie is great!")
} else {
  print("Not a favorite.")
}
```


**Multiple tests with `if…, else if…, else`**

Example 3
```{r}
x <- 0

if (x > 0) {
  print("x is positive")
} else if (x < 0) {
  print("x is negative")
} else {
  print("x is zero")
}
```

Example 4
```{r}
temperature <- 25

if (temperature >= 30) {
  weather <- "Hot"
} else if (temperature >= 20) {
  weather <- "Cool"
} else if (temperature >= 10) {
  weather <- "Breezy"
} else {
  weather <- "Freezing"
}

print(weather)
```


**Nested `if... else` statements**

Example 5 (note the modulo operator `%%` returns the remainder after a division)
```{r}
x <- 15

if (x > 0) {
  if (x %% 2 == 0) {
    print("x is a positive even number")
  } else {
    print("x is a positive odd number")
  }
} else {
  if (x %% 2 == 0) {
    print("x is a negative even number")
  } else {
    print("x is a negative odd number")
  }
}
```

**Vectorized tests with `ifelse()`**

Example 6
```{r}

ifelse(sqrt(16) > 3, 
       sqrt(16),
       0)
```


Example 7
```{r}
numbers <- c(10, 6, 7)

ifelse(numbers %% 2 == 1, 
       "odd",
       "even")
```


