---
title: "conditional statement"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TASK: Categorize penguins in three categories based on their body mass, and analyze the distribution of different categories.

Load the packages:
```{r}
library(palmerpenguins)
library(tidyverse)
```

Your Goal: 

Create a new column `mass_category` in the penguins dataset to categorize penguins based on their body mass, using the variable `body_mass_g` variable. If `body_mass_g` is:
* greater than 5000, label it as "Large"
* between 4000 and 5000 (inclusive), label it as "Medium"
* less than 4000, label it as "Small"

Then use `group_by` to group the data by `mass_category` and `summarize()` the average flipper length (variable `flipper_length_mm`) for each mass category.


## Answer the prompt using functions we already know: `case_when()`

Use `dplyr` functions `mutate()` and `case_when()` to create the new `mass_category` variable. Once ou are done, rewise the code to add an additional condition to check for missing data in `body_mass_g`, and set them to NA as `NA_Character_` in the new `mass_category` variable

```{r}

```


## Answer the prompt using new concepts from today: `ifelse()` function

Remember the syntax from the slides:
```
ifelse (condition to be evaluated,
    action performed when condition is TRUE,
    action performed when condition is FALSE)
```

```{r}

```



## Answer the prompt using using new concepts from today: write a set of `if-else` statements

In this case, at the top of your code you need to use `rowwise()` to make sure your if and else statements apply to a dataframe or, more specifically, to each of single row of the dataframe. In practice, `rowwise()` will force R to process each row individually (e.g., to consider each individual row as its own group). The code works like that:

```{r}
penguins %>%
  rowwise() %>%
  mutate(
    # add your if and else statements here
  )
```


```{r}

```
