---
title: "Conditional statements"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TASK: Categorize penguins in three categories based on their body mass, and analyze the distribution of different categories.

Load the packages:
```{r}
library(palmerpenguins)
library(tidyverse)
```

Your Goal: 

Create a new column `mass_category` in the penguins dataset to categorize penguins based on their body mass, using the variable `body_mass_g` variable. If `body_mass_g` is:
* greater than 5000, label it as "Large"
* between 4000 and 5000 (inclusive), label it as "Medium"
* less than 4000, label it as "Small"

Then use `group_by` to group the data by `mass_category` and `summarize()` the average flipper length (variable `flipper_length_mm`) for each mass category.


## Answer the prompt using functions we already know: `case_when()`

Use `dplyr` functions `mutate()` and `case_when()` to create the new `mass_category` variable. Once ou are done, rewise the code to add an additional condition to check for missing data in `body_mass_g`, and set them to NA as `NA_Character_` in the new `mass_category` variable
```{r}
penguins %>%
  mutate(
    mass_category = case_when(
      is.na(body_mass_g) ~ NA_character_,             
      body_mass_g > 5000 ~ "Large",
      body_mass_g >= 4000 & body_mass_g <= 5000 ~ "Medium",
      body_mass_g < 4000 ~ "Small"
    )
  ) %>%
  group_by(mass_category) %>%
  summarize(
    avg_flipper_length = mean(flipper_length_mm, na.rm = TRUE)
  )
```


## Answer the prompt using new concepts from today: `ifelse()` function

Remember the syntax from the slides:
```
ifelse (condition to be evaluated,
    action performed when condition is TRUE,
    action performed when condition is FALSE)
```

Solution without checking for NA:
```{r}
penguins %>%
  mutate(mass_category = "Small") %>% # count(mass_category)
  mutate(mass_category = ifelse(body_mass_g >= 4000 & body_mass_g <= 5000, 
                                "Medium",
                                mass_category)) %>%
  mutate(mass_category = ifelse(body_mass_g > 5000, 
                                "Large",
                                mass_category)) %>%
  group_by(mass_category) %>%
  summarize(avg_flipper_length = mean(flipper_length_mm, na.rm = TRUE))
```

Solution checking for NA:
```{r}
penguins %>%
  mutate(mass_category = ifelse(is.na(body_mass_g), 
                                NA_character_, 
                                "Small")) %>%
  mutate(mass_category = ifelse(body_mass_g >= 4000 & body_mass_g <= 5000, 
                                "Medium", 
                                mass_category)) %>%
  mutate(mass_category = ifelse(body_mass_g > 5000, 
                                "Large", 
                                mass_category)) %>%
  group_by(mass_category) %>%
  summarize(avg_flipper_length = mean(flipper_length_mm, na.rm = TRUE))
```

Solution with nested `if_else()` (NOT recommended!):
```{r}
penguins %>%
  mutate(mass_category = ifelse(body_mass_g > 5000, "Large",
                                ifelse(body_mass_g >= 4000, "Medium", "Small"))) %>%
  group_by(mass_category) %>%
  summarize(avg_flipper_length = mean(flipper_length_mm, na.rm = TRUE))
```



## Answer the prompt using using new concepts from today: write a set of `if-else` statements

In this case, at the top of your code you need to use `rowwise()` to make sure your if and else statements apply to a dataframe or, more specifically, to each of single row of the dataframe. In practice, `rowwise()` will force R to process each row individually (e.g., to consider each individual row as its own group). The code works like that:
```{r}
penguins %>%
  rowwise() %>%
  mutate(
    # add your if and else statements here
  )
```

Solution (must include check for NA at the top):
```{r}
penguins %>%
  rowwise() %>%
  mutate(
    mass_category = if (is.na(body_mass_g)) {
      NA_character_
    } else if (body_mass_g > 5000) {
      "Large"
    } else if (body_mass_g >= 4000 & body_mass_g <= 5000) {
      "Medium"
    } else {
      "Small"
    }
  ) %>%
  ungroup() %>%
  group_by(mass_category) %>%
  summarize(
    avg_flipper_length = mean(flipper_length_mm, na.rm = TRUE)
  )
```

## KEY DIFFERENCE BETWEEN THESE APPROACHES:

* `case_when()` and `ifelse()`: vectorized, meaning they evaluate a given condition over all elements (rows) of an entire column or variable (e.g. a vector) at the same time!

* if / else statements: not vectorized, meaning they evaluate a given condition one element at a time! So to apply them to a dataframe, you need to use a for loop or other tricks (such as `rowwise()`, like we did here) to evaluate each element (row) of the variable individually


## TAKE HOME:

1. `case_when()` and `ifelse()` have similar uses: to manipulate variables (recode, categorize, etc.) based on conditions inside dataframes; use them inside dataframes and with tidyverse!

2. Use if / else only outside dataframes and for so-called "flow control" operations (e.g., code executes based on conditions), the slides for today include examples of standard use of if / else statements 